Dish: Distribution-oblivious Shell Scripting

### 2019-11-28 pldi/draft.tex
Update section 3

% FIXME: \nv{TODO: remember to bubble up(me)}
An important characteristic of the \unix shell is its language-agnostic extensibility~\sx{bg}.
Users composing pipelines are free to install custom commands available from various sources.
Being general programs, commands are developed in a variety of languages and are extended over long periods of time.
%% \kk{Text point: This restricts the space of solutions for
%%   distribution of shell scripts}
As a result, any static characterization of shell commands is unsatisfactory as it would quickly become obsolete, supporting only a small subset of commands.
This poses a challenge to the developers of individual commands---how can they augment their commands to capture information critical to \sys without too much effort?
%% \kk{Text point: The characteristics of a good solution}

To address additions and changes to the set of commands available to the shell, \sys provides command developers a lightweight, domain-specific language (DSL).
The DSL allows developers to express semantic information about a command's distributanility that is critical to \sys.
This DSL can be used by the developers of new commands to succinctly express this information, such the class of a command and the sequence of sources from which it reads its input.
It could also be used by developers that extend or maintain existing commands, to express additions or changes to the command's implementation or interface.

@@

As simple examples, consider \ttt{cat} and \ttt{sort}.
By default \ttt{cat} is in \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter.
Conversely, \ttt{sort} defaults to \pur, but \ttt{with} \ttt{-h} it jumps into \sta.
% TODO: closure?

@@

\heading{Command Inputs and Outputs}
%
Consider the command \ttt{comm}. It takes as input two files and
performs a join operation. When invoked without any option, it
produces three-column output. The first column contains lines that
unique to the first file, the second column contains lines that are
unique to the second file, and the last column contains lines that
exist in both files. In the general case \ttt{comm} is in \pur, since
it needs to wait until it reads both files completely before it can
output its results.

However, \ttt{comm} can also be invoked using some options that
suppress the output of any combination of columns. More precisely, the
option \ttt{-1} suppresses the first column of the output. Similarly,
\ttt{-2} and \ttt{-3} supress the second and third column. Using both
\ttt{-2} (or \ttt{-1}) and \ttt{-3}, the command becomes \sta
regarding its first (resp. the second) file argument.

@@

\heading{Distributability DSL}
%
\sys's distributability description DSL (Fig.~\ref{fig:dsl}) is
designed to be able to classify a commands into multiple categories
given their options.

It assumes that the arguments of a command can always be separated to
options and file arguments (checking if they start with a
\ttt{-}~\footnote{Note that the single character \ttt{-} is not an
  option as it is most often used to designate that a command reads
  from its standard input}). A file written in the language contains a
record containing a mapping from predicates to assignments for each
command. Predicates can be simple boolean formulas on the existence
and the value of any option. Each assignment contains the command
category, the sequence of its input sources, and its output
source. Inputs to commands are a sequence of file arguments and its
standard input. Similarly, the output is either its standard output or
a file argument. An example record for \ttt{comm} can be seen below:

@@

%% \heading{Pure Commands}
%
\noindent
In addition to the commands in \sta that can be parallelized in a
straightforward way, there also exists a subclass of commands in \pur
that can be parallelized using divide-and-conquer techniques (as
mentioned in \cref{cmd}). In order to support the parallelization of
arbitrary commands in this category, \sys allows command developers to
provide a custom map and reduce step for these commands, where the map
is in \sta and the reduce in \pur, and their composition produces the
same output with the initial command. \sys can then replace occurences
of these commands with their map and reduce parts, enabling their
parallelization.

>>>>

% FIXME: \nv{TODO: remember to bubble up(me)}
An important characteristic of the \unix shell is its language-agnostic extensibility~\sx{bg}.
Users composing pipelines are free to install custom commands available from various sources.
Being general programs, commands are developed in a variety of languages and are extended over long periods of time.
%% \kk{Text point: This restricts the space of solutions for
%%   distribution of shell scripts}
As a result, any static characterization of shell commands is unsatisfactory and would quickly become obsolete, supporting only a small subset of the commands needed in practice.
This poses a challenge to the developers of individual commands---how can they augment their commands to capture information critical to \sys without too much effort?
%% \kk{Text point: The characteristics of a good solution}

To address % additions and changes to the set of commands available to the shell,
  this challenge, \sys arms command developers with a lightweight, domain-specific language (DSL).
The DSL allows developers to express semantic information about a command's distributability that is critical to \sys.
It can be used by the developers of new commands to succinctly express this information, such the class of a command and the sequence of sources from which it reads its input.
It can also be used by maintainers extending or maintain existing commands, to express additions or changes to the command's implementation or interface.

@@

As simple examples, consider \ttt{cat} and \ttt{sort}.
By default \ttt{cat} is in \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter.
Conversely, \ttt{sort} defaults to \pur, but \ttt{with} \ttt{-h} it jumps into \sta because it generates a static stream regardless of other input.
% TODO: closure?

@@

\heading{Command Inputs and Outputs}
A second concern captured by the DSL is the order of inputs and outputs.
Commands generally read multiple input streams (and, at cases, write to multiple output streams).
The invariants between inputs and outputs must be maintained over the distributed execution of the program.
% but these maintain different invariants, enabling very different distributability characteristics.

Consider the command \ttt{comm} that takes as input two files and
performs a join operation. When invoked without any option, it
produces a three-column output. The first column contains lines that
are unique to the first file, the second column contains lines that are
unique to the second file, and the last column contains lines that
exist in both files. In the general case \ttt{comm} is in \pur, since
it needs to wait until it reads both files completely before it can
output its results.

However, \ttt{comm} can also be invoked using some options that
suppress the output of any combination of columns. More precisely, the
option \ttt{-1} suppresses the first column of the output. Similarly,
\ttt{-2} and \ttt{-3} supress the second and third column. Using both
\ttt{-2} (or \ttt{-1}) and \ttt{-3}, the command jumps to \sta
regarding its first (resp. the second) file argument.

@@

\heading{Distributability DSL}
%
\sys's distributability description DSL (Fig.~\ref{fig:dsl}) is
designed to classify commands into multiple categories
given their options.

It assumes that the arguments of a command can always be separated to
options and file arguments (checking if they start with a
\ttt{-}~\footnote{Note that the single character \ttt{-} is not an
  option as it is most often used to designate that a command reads
  from its standard input}). A file written in the DSL contains a
record containing a mapping from predicates to assignments for each
command. Predicates can be simple boolean formulas on the existence
and the value of any option. Each assignment contains the command
category, the sequence of its input sources, and its output
source. Inputs to commands are a sequence of file identifiers and its
standard input. Similarly, the output is either its standard output or
a file identifier. An example record for \ttt{comm} can be seen below:

@@

%% \heading{Pure Commands}
%
\noindent
In addition to the commands in \sta that can be parallelized in a
straightforward way, there also exists a subclass of commands in \pur
that can be parallelized using divide-and-conquer techniques (as
mentioned in \cref{cmd}). In order to support the parallelization of
arbitrary commands in this category, \sys allows command developers to
provide custom map and reduce steps for them, where the map
is in \sta and the reduce in \pur, and their composition produces the
same output with the initial command. \sys can then replace occurrences
of these commands with their map and reduce parts, enabling their
parallelization.
